---
title: "607-wk6-project2-NZbizdata"
author: "Tyler Frankenberg"
date: "3/14/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Introduction

This data comes from a biennial survey of New Zealand businesses, on what percentage of their activity is conducted via the internet.

It is available from: <http://infoshare.stats.govt.nz/ViewTable.aspx?pxID=12482052-c8dd-47f1-9a2e-36fb3851b3fd>.

## Import Data

We start by importing the original table from its .csv file.  We include null value handling for blank values as well as for the ".." in non-survey years.

We have quite a mess on our hands...

```{r import}

nzbizdata_raw <- as_tibble(read_csv("https://raw.githubusercontent.com/curdferguson/607-wk5-hw5/master/nzbizdata.csv", col_names = FALSE, na=c("", "..")))

nzbizdata_raw

```

## Remove non-survey years

It's immediately clear this is a biennial survey taking place only in even-numbered years, so We'll start by removing rows that correspond to non-survey years.

```{r drop-rows}

nzbizdata_raw <- nzbizdata_raw %>% filter(is.na(`X2`) == FALSE)

nzbizdata_raw

``` 

## Transpose the tibble

There are two variables, which we'll call "Business Size" and "Percentage", stored in rows 1-2.

Rather than try to pivot_longer with multiple variables, we'll simplify the process by transposing the table, and then pivoting on a single variable, "Years".

```{r transpose}

# Transpose rows and columns
nzbizdata_transpose <- t(nzbizdata_raw)

nzbizdata_transpose

``` 
## Clean up our transposed data

This transposed matrix is a structure we'll be able to pivot much easier, after a little clean up...

```{r re-tibble}

# Fill in missing column headers in row 1, and make these the column headers
nzbizdata_transpose[1, c(1:2)] <- c("Business Size", "Percentage")
colnames(nzbizdata_transpose) <- nzbizdata_transpose[1,]

# Convert the data back to a tibble and drop the now-rendundant "header" row
nzbizdata_transpose <- as_tibble(nzbizdata_transpose)
nzbizdata <- nzbizdata_transpose %>% slice(2:25)

# Convert numerical columns to data type "int"
nzbizdata[, c(3:8)] <- nzbizdata[, c(3:8)] %>% type.convert()

nzbizdata

``` 

# Get Values for Business Size

This tibble will be much better for pivoting.  However, we still have missing values to take care of in our "Business Size" variable.

```{r biz_size}

# Create a vector of the unique values for "Business Size"
biz_size <- as.vector(nzbizdata$`Business Size`)
biz_size <- unique(biz_size[!is.na(biz_size)])

# Replace NA values with the appropriate value
nzbizdata[c(2:6), 1] <- biz_size[[1]]
nzbizdata[c(8:12), 1] <- biz_size[[2]]
nzbizdata[c(14:18), 1] <- biz_size[[3]]
nzbizdata[c(20:24), 1] <- biz_size[[4]]

# Convert "Business Size" to factor format so we can sort it appropriately
nzbizdata$`Business Size` <- nzbizdata$`Business Size` %>% factor(levels = biz_size)

nzbizdata

```
## Pivot

What we're working with now is an easy pivot to a "long" format, using pivot_longer:

```{r pivot}

# Create a vector for years
years <- colnames(nzbizdata[, c(3:8)])

# Pivot the data and sort by year
nzbizdata_long <- nzbizdata %>% 
  pivot_longer(years, names_to="Year", values_to="Count") %>%
  arrange(`Year`, `Business Size`)

nzbizdata_long

```


